FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-20 NewVersion/paper.tex

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,override},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\sffamily\bfseries,
	aboveskip=0pt,
	belowskip=0pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

>>>>

\lstset{ %
	language=Java,                % choose the language of the code
	columns=flexible,
	lineskip=-1pt,
	basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
	numbers=none,                   % where to put the line-numbers
	numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
	stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
	numbersep=5pt,                  % how far the line-numbers are from the code
	backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
	showspaces=false,               % show spaces adding particular underscores
	showstringspaces=false,         % underline spaces within strings
	showtabs=false,                 % show tabs within strings adding particular underscores
	morekeywords={var,override},
	%  frame=single,                   % adds a frame around the code
	tabsize=2,                  % sets default tabsize to 2 spaces
	captionpos=none,                   % sets the caption-position to bottom
	breaklines=true,                % sets automatic line breaking
	breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
	title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
	escapeinside={(*}{*)},          % if you want to add a comment within your code
	keywordstyle=\sffamily\bfseries,
	aboveskip=3pt,
	belowskip=3pt
	% commentstyle=\color{Gray},
	% stringstyle=\color{Green}
}

### 2017-10-20 NewVersion/sections/Auxiliary.tex
small fix. to 25pages
added missing figure and small fix

\begin{flalign*}
	& \rhd \textit{Definition of } \mbody(m, I_d, I_s): & \\
	& \bullet \mbody(m, I_d, I_s) = (J, \overline{I_x} \; \overline{x}, I_e \; e_0) & \\
	& \indent\indent \textrm{with: } \mostSpecific(m, I_d, I_s) = \{I\} & \\
	& \hspace{.77in} \mostSpecificOverride(m, I_d, I) = \{J\} & \\
	& \hspace{.77in} J[m\ \kwoverride\ I] = \method{I_e}{m}{I_x}{x}{I}{e_0} & \\
	& \bullet \mbody(m, I_d, I_s) = (J, \overline{I_x} \; \overline{x}, I_e \; \o) & \\
	& \indent\indent \textrm{with: } \mostSpecific(m, I_d, I_s) = \{I\} & \\
	& \hspace{.77in} \mostSpecificOverride(m, I_d, I) = \{J\} & \\
	& \hspace{.77in} J[m\ \kwoverride\ I] = \absmethod{I_e}{m}{I_x}{x}{I} & \\
\end{flalign*}
To calculate $\mbody(m, I_d, I_s)$, the invocation of $\mostSpecific$ looks for the most specific original methods and their interfaces, and expects a singleton set, so as to avoid unambiguity. Furthermore, the invocation of $\mostSpecificOverride$ also expects a unique (unambiguous) most specific hierarchical override. And finally the target method is returned.

@@

\begin{flalign*}
	& \rhd \textit{Definition of } \mostSpecific(m, I, J): & \\
	& \bullet \mostSpecific(m, I, J) = \prune(origins) & \\
	& \indent\indent \textrm{with: } origins = \{K \mid \subt{I}{K}, \textrm{ and } \subt{K}{J} \; \lor \; \subt{J}{K}, &\\
	& \hspace{1.62in} \textrm{ and } K[m\ \kwoverride\ K] \textrm{ is defined} \} &
\end{flalign*}
By the definition, an interface belongs to $\mostSpecific(m, I, J)$ if and only if:
\begin{itemize}
	\item It originally defines $m$;
	\item It is a super type of $I$ (including $I$);
	\item It is either a super type or a subtype of $J$ (including $J$);
	\item Any subtype of it does not belong to the same result set because of $\prune$.
\end{itemize}

@@

\begin{flalign*}
	& \rhd \textit{Definition of } \mostSpecificOverride(m, I, J): & \\
	& \bullet \mostSpecificOverride(m, I, J) = \prune(overrides) & \\
	& \indent\indent \textrm{with: } overrides = \{K \mid \subt{I}{K}, \; \subt{K}{J} \textrm{ and } K[m\ \kwoverride\ J] \textrm{ is defined} &
\end{flalign*}
By the definition, an interface belongs to $\mostSpecificOverride(m, I, J)$ if and only if:
\begin{itemize}
	\item it is between $I$ and $J$ (including $I$, $J$);
	\item it hierarchically overrides $J.m$;
	\item any subtype of it does not belong to the same set.
\end{itemize}

@@

%%%%============================ prune(set) ================%%%%%%%%
\begin{flalign*}
	& \rhd \textit{Definition of } \prune(set): & \\
	& \bullet \prune(set) = \{I \in set \; | \; \nexists J \in set\setminus I, J <: I\} &
\end{flalign*}
The $\prune$ function takes a set of
types, and filters out those that have subtypes in the same set. In the returned set,
none of them has subtyping relation to one another, since all super types have been removed.
%%%%============================ prune(set) end ================%%%%%%%%

%%%%============================ canOverride ================%%%%%%%%
\begin{flalign*}
	& \rhd \textit{Definition of } \canOverride(m, I, J): & \\
	& \bullet \canOverride(m, I, J) = True & \\
	& \indent\indent \textrm{with: } I[m\ \kwoverride\ I] = I_e \; m(\overline{I_x} \; \overline{x}) \; \kwoverride \; I \ldots & \\
	& \hspace{.77in} J[m\ \kwoverride\ J] = I_e \; m(\overline{I_x} \; \overline{y}) \; \kwoverride \; J \ldots &
\end{flalign*}
$\canOverride$ just checks that two original $m$ in $I$ and $J$ have the same type.
%%%%============================ canOverride end ================%%%%%%%%

%%%%============================ canInstantiate ================%%%%%%%%
\begin{flalign*}
	& \rhd \textit{Definition of } \canInstantiate(I): & \\
	& \bullet \canInstantiate(I) = True & \\
	& \indent\indent \textrm{with: } \forall m, \forall J \in \mostSpecific(m, I, I), \mostSpecificOverride(m, I, J) = \{K\}, & \\
	& \hspace{.77in} \textrm{ and } K[m\ \kwoverride\ J] = \method{I_e}{m}{I_x}{x}{J}{e_0} &
\end{flalign*}
$\canInstantiate(I)$ checks whether interface $I$ can be instantiated by the keyword $\kwnew$.
$\mostSpecific(m, I, I)$ represents the set of branches $I$ inherits on method $m$. $I$ can be instantiated
if and only if for every branch, the most specific implementation is unambiguous and non-abstract.
%%%%============================ canInstantiate end ================%%%%%%%%

>>>>

\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \mbody(m, I_d, I_s): & \\
	& \bullet \mbody(m, I_d, I_s) = (J, \overline{I_x} \; \overline{x}, I_e \; e_0) & \\
	& \indent\indent \textrm{with: } \mostSpecific(m, I_d, I_s) = \{I\} & \\
	& \hspace{.77in} \mostSpecificOverride(m, I_d, I) = \{J\} & \\
	& \hspace{.77in} J[m\ \kwoverride\ I] = \method{I_e}{m}{I_x}{x}{I}{e_0} & \\
	& \bullet \mbody(m, I_d, I_s) = (J, \overline{I_x} \; \overline{x}, I_e \; \o) & \\
	& \indent\indent \textrm{with: } \mostSpecific(m, I_d, I_s) = \{I\} & \\
	& \hspace{.77in} \mostSpecificOverride(m, I_d, I) = \{J\} & \\
	& \hspace{.77in} J[m\ \kwoverride\ I] = \absmethod{I_e}{m}{I_x}{x}{I} & \\
\end{flalign*}
\saveSpaceFig

To calculate $\mbody(m, I_d, I_s)$, the invocation of $\mostSpecific$ looks for the most specific original methods and their interfaces, and expects a singleton set, so as to avoid unambiguity. Furthermore, the invocation of $\mostSpecificOverride$ also expects a unique (unambiguous) most specific hierarchical override. And finally the target method is returned.

\begin{figure*}[t]
\saveSpaceFig
\begin{mathpar}
	% \sinvk \\
	\spathinvk \\
	% \ssuperinvk \\
	% \sstaticinvk \\
	\creceiver \hspace{.5in}
	% \cpathreceiver \\
	\cargs \\
	% \cpathargs \\
	% \csuperargs \\
	% \cstaticargs \\
	\cstatictype \\
	\cfreduce \\
	\cannoreduce
\end{mathpar}
\caption{Small-step semantics.}\label{fig:smallstep}
\saveSpaceFig
\end{figure*}

@@

\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \mostSpecific(m, I, J): & \\
	& \bullet \mostSpecific(m, I, J) = \prune(origins) & \\
	& \indent\indent \textrm{with: } origins = \{K \mid \subt{I}{K}, \textrm{ and } \subt{K}{J} \; \lor \; \subt{J}{K}, &\\
	& \hspace{1.62in} \textrm{ and } K[m\ \kwoverride\ K] \textrm{ is defined} \} &
\end{flalign*}
By the definition, an interface belongs to $\mostSpecific(m, I, J)$ if and only if:
\begin{itemize}
	\item It originally defines $m$;
	\item It is a super type of $I$ (including $I$);
	\item It is either a super type or a subtype of $J$ (including $J$);
	\item Any subtype of it does not belong to the same result set because of $\prune$.
\end{itemize}
\saveSpaceFig

@@

\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \mostSpecificOverride(m, I, J): & \\
	& \bullet \mostSpecificOverride(m, I, J) = \prune(overrides) & \\
	& \indent\indent \textrm{with: } overrides = \{K \mid \subt{I}{K}, \; \subt{K}{J} \textrm{ and } K[m\ \kwoverride\ J] \textrm{ is defined} &
\end{flalign*}
\saveSpaceFig

By the definition, an interface belongs to $\mostSpecificOverride(m, I, J)$ if and only if:
\begin{itemize}
	\item it is between $I$ and $J$ (including $I$, $J$);
	\item it hierarchically overrides $J.m$;
	\item any subtype of it does not belong to the same set.
\end{itemize}

@@

%%%%============================ prune(set) ================%%%%%%%%
\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \prune(set): & \\
	& \bullet \prune(set) = \{I \in set \; | \; \nexists J \in set\setminus I, J <: I\} &
\end{flalign*}

The $\prune$ function takes a set of
types, and filters out those that have subtypes in the same set. In the returned set,
none of them has subtyping relation to one another, since all super types have been removed.
%%%%============================ prune(set) end ================%%%%%%%%

%%%%============================ canOverride ================%%%%%%%%
\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \canOverride(m, I, J): & \\
	& \bullet \canOverride(m, I, J) = True & \\
	& \indent\indent \textrm{with: } I[m\ \kwoverride\ I] = I_e \; m(\overline{I_x} \; \overline{x}) \; \kwoverride \; I \ldots & \\
	& \hspace{.77in} J[m\ \kwoverride\ J] = I_e \; m(\overline{I_x} \; \overline{y}) \; \kwoverride \; J \ldots &
\end{flalign*}
\saveSpaceFig
$\canOverride$ just checks that two original $m$ in $I$ and $J$ have the same type.
%%%%============================ canOverride end ================%%%%%%%%

%%%%============================ canInstantiate ================%%%%%%%%
\saveSpaceFig
\begin{flalign*}
	& \rhd \textit{Definition of } \canInstantiate(I): & \\
	& \bullet \canInstantiate(I) = True & \\
	& \indent\indent \textrm{with: } \forall m, \forall J \in \mostSpecific(m, I, I), \mostSpecificOverride(m, I, J) = \{K\}, & \\
	& \hspace{.77in} \textrm{ and } K[m\ \kwoverride\ J] = \method{I_e}{m}{I_x}{x}{J}{e_0} &
\end{flalign*}

$\canInstantiate(I)$ checks whether interface $I$ can be instantiated by the keyword $\kwnew$.
$\mostSpecific(m, I, I)$ represents the set of branches $I$ inherits on method $m$. $I$ can be instantiated
if and only if for every branch, the most specific implementation is unambiguous and non-abstract.
%%%%============================ canInstantiate end ================%%%%%%%%

### 2017-10-20 NewVersion/sections/Discussion.tex

\begin{figure*}[t]
    % \nocaptionrule
    % \centering
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=2cm]{pics/P7.pdf}
    \end{minipage}
    \centering
    \hspace*{2pt}
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=2cm]{pics/P8.pdf}
    \end{minipage}  
    \caption{Triangle inheritance (left) and diamond inheritance (right) on abstract methods.}\label{fig:abstractdiamond}
\end{figure*}

\subsection{Orthogonal \& Non-Orthogonal Extensions}\label{sec:orthoext}

Our model is designed as a minimal calculus for focusing only on resolving unintentional conflicts. Therefore, we have omitted a number of
common orthogonal features including primitive types, assignments, method overloading, covariant method return types, static dispatch, and so on.
Those features can, in principle, be modularly added to the model without breaking type soundness. For example, we present the additional syntax, typing and semantic rules of static invocation below as an extension:

@@

\MIM{} rejects the following case of diamond inheritance:

\vspace{3pt}\begin{lstlisting}
interface A { void m() {...} }
interface B extends A { void m() {...} }
interface C extends A { void m() {...} }
interface D extends B, C {}
\end{lstlisting}\vspace{3pt}
Here both $B.m$ and $C.m$ override $A.m$, and D inherits
both conflicting methods without an explicit override. In this case automatically merging 
the two methods (to achieve diamond inheritance) is not possible, which is why many 
models (like traits and Java 8) reject such programs. Moreover, keeping the two 
method implementations in \lstinline|D| is problematic. 
In essence, hierarchical information is 
not helpful to disambiguate later method calls, since the two methods share 
the same origin (\lstinline|A.m|). 
Our calculus rejects such conflicts by the \textsc{(T-Intf)} rule, where $D$ is considered to be ill-formed. 
We believe that rejecting \lstinline|D| follows the principle of models like traits and Java 8 interfaces, 
where the language/type-system is meant to alert the programmer for a possible conflict early.

>>>>

\begin{figure*}[t]
\saveSpaceFig
    % \nocaptionrule
    % \centering
    \centering
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=1.4cm]{pics/P7.pdf}
    \end{minipage}
    \centering
    \hspace*{2pt}
    \begin{minipage}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=1.2cm]{pics/P8.pdf}
    \end{minipage}  
    \caption{Triangle inheritance (left) and diamond inheritance (right) on abstract methods.}\label{fig:abstractdiamond}
\saveSpaceFig
\end{figure*}

\subsection{Orthogonal \& Non-Orthogonal Extensions}\label{sec:orthoext}
Our model is designed as a minimal calculus for focusing only on resolving unintentional conflicts. Therefore, we have omitted a number of
common orthogonal features including primitive types, assignments, method overloading, covariant method return types, static dispatch, and so on.
Those features can, in principle, be modularly added to the model without breaking type soundness. For example, we present the additional syntax, typing and semantic rules of static invocation below as an extension:

@@

\MIM{} rejects the following case of diamond inheritance:
\begin{lstlisting}
interface A { void m() {...} }
interface B extends A { void m() {...} }
interface C extends A { void m() {...} }
interface D extends B, C {}
\end{lstlisting}
Here both $B.m$ and $C.m$ override $A.m$, and D inherits
both conflicting methods without an explicit override. In this case automatically merging 
the two methods (to achieve diamond inheritance) is not possible, which is why many 
models (like traits and Java 8) reject such programs. Moreover, keeping the two 
method implementations in \lstinline|D| is problematic. 
In essence, hierarchical information is 
not helpful to disambiguate later method calls, since the two methods share 
the same origin (\lstinline|A.m|). 
Our calculus rejects such conflicts by the \textsc{(T-Intf)} rule, where $D$ is considered to be ill-formed. 
We believe that rejecting \lstinline|D| follows the principle of models like traits and Java 8 interfaces, 
where the language/type-system is meant to alert the programmer for a possible conflict early.
